Inferring the gravitational potential of the Milky Way with a few precisely measured stars

### 2014-02-10 text/paper2/ms.tex
cleanup

The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles. The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the tidal radius with the isotropic, log-normal centered on the velocity dispersion of the satellite:
\begin{align}
	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}\\
	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}\label{eq:int} \\
	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\right]_{t=\tau}
\end{align}
% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$. We then compute the marginal likelihood, $p(\W \given \W_s, \Phi)$, by integrating over all possible unbinding times,
\begin{align}
	p(\W \given \W_s, \Phi) = \int p(\W \given \W_s, \tau, \Phi)p(\tau) d\tau.
\end{align}
In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the satellite and parent.

@@

\subsection{\gaia + \spitzer}
64 stars in Sgr-like stream

\subsection{\spitzer + current data}
64 stars in Sgr-like stream

\subsection{\spitzer + current data, no progenitor}
32 stars in Orphan-like stream, XX stars in GD-1 -like stream

@@

\acknowledgements
Binney, Rix, Sanders

>>>>

The likelihood of a star's position today is computed by first transforming the observed position to Galactocentric cartesian coordinates and integrating the orbits of the satellite and star backwards for several Gigayears in the parent potential, $\Phi$, treating both as test particles. The distance, $R$, and velocity, $V$, of the star relative to the satellite are computed at each time-step, and the likelihood is computed as the product of the isotropic, log-normal centered on the tidal radius with the isotropic, log-normal centered on the velocity dispersion of the satellite:
\begin{align}
	p(\W \given \W_s, \tau, \Phi) &= p(\X \given \X_s, \tau, \Phi) \jac{(l,b,D,\mu_l,\mu_b,v_r)}_{t=\tau}\\
	p(\X \given \X_s, \tau, \Phi) &= p(\ln R, \ln V \given \tau, \Phi) \jac{(x,y,z,v_x,v_y,v_z)}_{t=\tau}\label{eq:int} \\
	p(\ln R, \ln V \given \tau, \Phi) &= \left[\mathcal{N}(\ln R | \ln \rtide)\mathcal{N}(\ln V | \ln \sigma_v)\right]_{t=\tau}
\end{align}
% , but the vectors $\D$ and $\W$ [exist] in different coordinate systems --- heliocentric spherical and Galactocentric cartesian, respectively --- so we must also track the Jacobian that [describes] this transformation:
where in eq.~\ref{eq:int} we have implicitly integrated over and neglected the two angle coordinates associated with $\ln R$ and the two angles associated with $\ln V$. We then compute the marginal likelihood, $p(\W \given \W_s, \Phi)$, by integrating over all possible unbinding times,
\begin{align}
	p(\W \given \W_s, \Phi) = \int p(\W \given \W_s, \tau, \Phi)p(\tau) d\tau.
\end{align}
In general, $p(\tau)$ is likely conditional on the orbit of the progenitor --- stars are preferentially stripped at pericentric passages --- but this is neglected and is taken to be uniform over the entire interaction history between the satellite and parent.

@@

\subsection{\gaia + \spitzer}
4 stars in Sgr-like stream

\subsection{\spitzer + current data}
4 stars in Sgr-like stream

\subsection{\spitzer + current data, no progenitor}
4 stars in Orphan-like stream, XX stars in GD-1 -like stream

@@

\acknowledgements
